Stack Safety as a Security Property

### 2020-07-02 main.tex

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code, either
directly (via a jump instruction) or indirectly (by corrupting the return
address on the stack and then returning). \leo{There needs to be some
  informal description of the wbcf part here. Sean?}

@@

\judgment{\(f\ \traceelem\)}
         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
%
\judgment{}
         {\(\PUT(f,\traceelem) = \traceelem\)}
%
\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. Conversely\bcp{why conversely?}, if the predicate never
holds, \(\PUT\) is the identity function.

@@

\judgmenttwo{\(\mach \hookrightarrow \machT\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT)
              = \machT'\)}
            {\(\mach \hookrightarrow \machT' | f\)}
\noindent
And similarly for traces with policies:

@@

\leo{TODO: We really need a better layout than centered paragraphs.}
\judgment{}{\(\obsT \lesssim \obsT\)}

\judgment{}{\(\tau \lesssim \obsT\)}
\judgment{}{\(w \lesssim w\obsT\)}

\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\tau \obsT_1 \lesssim \obsT_2\)}

\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\obsT_1 \lesssim \tau \obsT_2\)}

\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(w\obsT_1 \lesssim w\obsT_2\)}

@@

We then define equivalence\bcp{Can we please call it something else?  An
  ``equivalence'' should always be transitive.} of observation traces as traces prefixing each other:
\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]

@@

\definition{Variations}
We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another given a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
if it started from a variant initial machine state.

@@

\leo{Discuss variations in a similar manner}

To lift trace confidentiality to a system property, we follow the same
pattern as in stack integrity:

\sna{Version for stepping from a state}
\definition{Confidential Execution}
  Stack confidentiality, below, is a property of a system that, from a call to
  a return, each trace obeys a local confidentiality property,
  {\it confidential execution}.

  We will say that a state \(\mpstate{\mach}{\pol}\) satisfies confidential
  execution with regard to a contour \(\contour\) and an ending condition
  \(R\) that is a predicate on machine states if, for all variant states
  \(\mach' \approx_\contour \mach\), the induced traces from \(\mach\) and
  \(\mach'\) behave the same across three cases. Let
  \(\mpstate{\mach}{\pol} \hookrightarrow \MPT | R\) and
  \(\mach' \hookrightarrow \machT | R\).

  If \(\MPT\) ends with a state \(\mach_r\) and \(R\ \mach_r\), then \(\machT\)
  ends with a state \(\mach_r'\) such that \(R\ \mach_r'\). \(\MPT\) and
  \(\machT\) are observationally equivalent, and for each component
  \(\component\) such that \(\contour(\component) = \HIGHSEC\),
  \(\mach_r(\component) = \mach_r'(\component)\).

  If \(\MPT\) is infinite, then \(\machT\) is infinite, and they are
  observationally equivalent.

  If \(\MPT\) ends before \(R\) ever holds, meaning that it halted due to a
  policy violation, then it is a strict observational prefix of \(\machT\),
  which must be infinite.

@@

      Intuitively, the result of a rollback is an idealized state that
      reflects changes to accessible components, but ignores the callee's
      interference with the caller's data. Then the observable behavior of the
      resulting trace is the standard to which we expect the real trace to
      conform.

      A system enjoys observable integrity with regard to \(\callmap\) and
      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
      \(\pol_e\) in the trace induced by
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      with the return state \(\mach_r\) at the end of \(\MPT\), then the
      following holds.

      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
      levels of the call and
      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
      possible policy failure: \(\MPT \sim_P \machT'\).

    \definition{Observable Confidentiality}
      Observable confidentiality is concerned with whether high confidentiality
      data influences observable both behavior during and after a call. As with
      eager confidentiality, we will compare the behavior of the trace induced
      by a state to another induced by its variant. But if the original returns,
      before continuing, we must reset the variant to conform with the original
      state.

      The \(\mathit{restore}\) function takes a contour and four states: the
      original call state, a variant call state, and corresponding return
      states, and returns another state. The resulting state restores those
      components that are high confidentiality to match the original call,
      unless they were changed during execution:
      \[\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')(\component) =
        \begin{cases}
          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_) \text{ and } \\
                  & \mach_c(\component) = \mach_r(\component) \text{ and }
                    \mach_c'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}\]

      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_c\) and
      \(\mach_c'\), respectively, with \(\mach_c \approx_\contour \mach_c'\),
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim_O
      \pi_o(\machT)\).

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_c,\mach_c',\mach_r,\mach_r')
      \hookrightarrow \machT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim_O \pi_o(\machT')\).

      \sna{Structured like this, we need to start with the call state, not
           the entry. That's because we need to know what to restore to.
           But we could also define restore in the stack confidentiality
           section, maybe?}

@@

>>>>

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establshes safety between
functions: all code is owned by a function, and all jumps between functions
are either calls that enter at a marked entry point or returns that enter
just after their associated call in the standard nested call structure.

%\leo{There needs to be some informal description
%  of the wbcf part here. Sean?}

@@

\judgment{\(f\ \traceelem\)}
         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
%
\judgment{}
         {\(\PUT(f,\traceelem) = \traceelem\)}
%
\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
And similarly for traces with policies:

@@

%\leo{TODO: We really need a better layout than centered paragraphs.}

\begin{minipage}{.3\textwidth}
  \judgment{}{\(\obsT \lesssim \obsT\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{}{\(\tau \lesssim \obsT\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{}{\(w \lesssim w\obsT\)}
\end{minipage}

\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\tau \obsT_1 \lesssim \obsT_2\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(\obsT_1 \lesssim \tau \obsT_2\)}
\end{minipage}
\begin{minipage}{.3\textwidth}
\judgment{\(\obsT_1 \lesssim \obsT_2\)}
         {\(w\obsT_1 \lesssim w\obsT_2\)}
\end{minipage}

@@

We then define similarity of observation traces as traces prefixing each other:

\bcp{Can we please call it something else?  An
  ``equivalence'' should always be transitive.}
\sna{Similarity seems less loaded?}
\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]

@@

\definition{Variations}
We say two machine states $\mach$ and $\mach'$ are {\em variations} of one
another given a contour $\contour$ if they agree on low
confidentiality data:
%
\[
\mach \approx_\contour \mach' \triangleq \forall \component .
\contour(\component) = \LOWSEC \rightarrow \mach(\component) =
\mach'(\component)
\]
%
Then, the trace version of confidentiality operates on two traces: the
``original'' one $MP \in \MPOTRACE$ that corresponds to a callee $f$
and a ``variant'' one $M \in \MTRACE$ that would corresponding to $f$
if it started from a variant initial machine state. An example in

@@

Figure \ref{fig:lazyex1} shows the initial states \(\mach_e\), left,
and \(\mach_e'\), right. Low confidentiality data in green is shared
between them, while high confidentiality data in red and yellow may differ.
After executing the code shown, the final states below have written different
values to address 42, represented by the black and white components.
It is allowed for address 40 to differ between the states due to the initial
variation, but address 42 changed differently between the traces, representing
a leak of data. Meanwhile \ref{fig:lazyex2}, while high confidentiality data
is read, it is overwritten before return, and the write at step 3 always writes
the same value. Yellow and red locations differ between variations only because
of the initial variation, and the rest are identical. So this example passes.

\sna{TODO: it would be nice for this to correspond at least roughly to
     the layout in the running example.}

\begin{figure}
  % H - Body
  \[
  \begin{array}{r|c|c|c|c}
    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
    \tracerow{2}{104}{\sw ~ 42 ~ r_0}{}{Store that value to 42}
    \tracerow{3}{108}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
  \end{array}
  \]

  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\
  ~\\
  ~\\
 \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{red}
      \memory{1}{black}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{yellow}%
      \memory{1}{white}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\

  \caption{Failing Trace Confidentiality}
  \label{fig:lazyex1}
\end{figure}
\leo{Discuss variations in a similar manner}

\begin{figure}
  % H - Body
  \[
  \begin{array}{r|c|c|c|c}
    \tracerow{1}{100}{\lw ~ r_0 ~ 40}{}{Load value at 40 to $r_0$}
    \tracerow{2}{104}{\sw ~ 42 ~ 0}{}{Store a constant value to 42}
    \tracerow{3}{108}{\lw ~ r_0 ~ 45}{}{Overwrite $r_0$ with accessible value}
    \tracerow{4}{112}{\jalr ~ \rra ~ \rra ~ 0}{}{Return}
  \end{array}
  \]

  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{red}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \MemoryLabel{10em}{1.5em}{40}
      \MemoryLabel{15em}{1.5em}{42}
      \memory{2}{gray}%
      \memory{4}{yellow}[{\makebox[0pt]{\(\HIGHSEC\)}}]%
      \memory{2}{green}[{\makebox[0pt]{\(\LOWSEC\)}}]%
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\
  ~\\
  ~\\
 \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{red}
      \memory{1}{black}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage}
  \begin{minipage}{.49\textwidth}
    \begin{center}
      \memory{2}{gray}%
      \memory{3}{yellow}%
      \memory{1}{black}
      \memory{2}{green}
      \memory{2}{gray}%
      \\
    \end{center}
  \end{minipage} \\

  \caption{Passing Confidentiality Despite Failing Integrity}
  \label{fig:lazyex2}
\end{figure}
To lift trace confidentiality to a system property, we follow the same
pattern as in stack integrity:

@@

      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
      agrees with \(\mach_e\) on every component the callee should not change,
      but with \(\mach_r\) on those components that it is permitted to change.
      So the observable behavior that follows is the idealized behavior to
      which we compare the trace from \(\mach_r\), which must prefix it (to
      account for premature halts due to policy violations.)
      %
      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

    \definition{Observable Stack Integrity}

      A system enjoys {\em observable integrity} with regard to \(\callmap\)
      and  \(\codemap\) if for any initial state \(\mach_0\) and each
      transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and
      contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

    \definition{Observable Trace Confidentiality}

      Trace confidentiality is a property of two traces whose initial machine
      states are variations, with two sub-properties. First, the behavioral
      condition -- that their observable behaviors prior to the end of the
      trace are equivalent -- reflects the intuition that secret values should
      not influence visible behavior. Second, we have the end condition that
      their final states (if any) are equal up to the original variation. The
      end condition over-approximates the extension of the behavioral condition
      to future execution.

      We introduce {\em observable trace confidentiality}, which extends the
      behavioral condition to apply to execution from the final state on.
      This requires restoring varied components to their original values,
      since after return they will once again be accessible legitimately.

      In Figure \ref{fig:lazyex1}, the final states of the original and variant
      differ, because a written value (in black or white) depended on the value
      of a high confidentiality location. So the resulting trace fails trace
      confidentiality. For observable trace confidentiality, we ask whether the
      differing values at address 42 cause any observable distinctions in
      future behavior. To do so, we must undo the variation of formerly high
      confidentiality locations, so that those in yellow match those in red,
      while preserving the updated values where data may have been leaked.

      The \(\mathit{restore}\) function takes a contour and four states: the
      original call state, a variant call state, and corresponding return
      states, and returns another state. The resulting state restores those
      components that are high confidentiality to match the original call,
      unless they were changed during execution:
      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')(\component) =
        \begin{cases}
          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_) \text{ and } \\
                  & \mach_e(\component) = \mach_r(\component) \text{ and }
                    \mach_e'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}\]

      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_e\) and
      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machT)\).

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).

      We write this \(\text{\sc ObsConf}_\contour ~ \MPT\).

@@

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with regard to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT\).
